---
title: "MachineLearningFinal_Coursera"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step1: Collect the Data

Download the required data from the shared links and read it into the R-Environment

```{r}
library(dplyr)
library(caret)
pml_training<-read.csv("pml-training.csv")
pml_testing<-read.csv("pml-testing.csv")
```

## Step2: Data Cleanup
Many columns are having multiple NA columns and No data in it. It will be good to remove them as they will not contribute towards the prediction. Create a function to clean both the data. If any variables has more than 90% of missing values or no value, just remove those columns. Also, timestamp , name, index are not relevent for prediction. 

```{r}
clean_data<-function(dat_inp){
  ## remove the mostly NA (more than NA's in a column) and empty columns with more than 90% cells
  dat_out<- dat_inp %>% select_if(!colSums(is.na(.)) >= .9*nrow(.)) %>% 
    select_if(!colSums(.=="") > .9*nrow(.)) 
  ## remove the date column, index and other columns which are not relevent for prediction
  dat_out<- dat_out %>% select(-grep("name|timestamp|window|X", colnames(.), value=F))
  ## return the cleaned data
  return(dat_out)
}

## get the cleaned training data
training<-clean_data(pml_training)

## get the cleaned testing data for real time
validating<-clean_data(pml_testing)
```

## Step3: Data Spliting intotraining and testing
Divide the data into two parts using the function from caret package. Kepp 75% data for traing and 25% for testing purpose 

```{r}
## get the 75% as training and the remaining as testing 
inTrain<-createDataPartition(y=training$classe,p=.75,list=FALSE)
## get the respective data
training<-training[inTrain,]
testing<-training[-inTrain,]
```

## Step4: Train the model
Using random forest method from caret package for classification purpose with 5-fold validation instead of 10-fold just for time efficiency purpose

```{r}
ctrl <- trainControl(method = "cv", number = 5, selectionFunction = "oneSE")
fitMod <- train(classe ~ ., data = training, method = "rf", trControl = ctrl)
```

## Step5: Model evaluation
Evaluate the model using the test data created at step 3. 
```{r}
# predict outcomes using testing data
pred <- predict(fitMod, testing)
```

## Step6: Performance Analysis
Evaluate the model using the test data created at step 3.
```{r}
# Show prediction result
conf_mtr <- confusionMatrix(testing$classe, pred)
print(conf_mtr)

## accuracy
accuracy<-conf_mtr$overall["Accuracy"]
accuracy
```

## Step7: Get the results for real time data
Evaluate the model using the test data created at step 3.
```{r}
pred_real<-predict(fitMod,validating)
pred_real
```
